Exploring probability distributions for bivariate temporal granularities

### 2019-12-02 main.Rmd
section headings changed

# Cyclic Time granularities {#sec:circular-gran-def}

@@

## Quasi-circular {#sec:circular-gran-def}

@@

## Aperiodic {#sec:circular-gran-def}

@@

## Fixed and Conditioned variables {#sec:synergy}

@@

## Measured variable

The measured variable which is the univariate observational units can be dicrete or continuous.

@@

### Empirical methods

@@

```{r exdot, fig.cap="10th, 20th, 30th, 40th and 50th quantiles of runs per over are drawn  across overs of the innings with no (facet 1), one (facet 2) and two (facet 3) dot balls per over. For all three cases, difference in runs between subsequent overs decrease.", fig.pos="htb"}

>>>>

# Cyclic Time granularities {#sec:cyclic-gran-def}

@@

## Quasi-circular {#sec:quasi-circular-gran-def}

@@

## Aperiodic {#sec:aperiodic-gran-def}

@@

## Synergy of fixed and conditioned variables {#sec:synergy}

@@

## Density estimation of measured variable

The measured variable can be dicrete or continuous. The probability distribution of these univariate measured variable could be studied by estimating the proability density function (pdf for continuous) or probability mass function (pmf for dicrete). Both pdf and pmf are concerned with point probabilities of the random variable. Suppose we have observations which are realizations of univariate random variables, $X_1, X_2, \dots, X_n \sim F$, where $F$ denotes an unknown cumulative distribution function and $X_1, X_2, \dots, X_n$ are i.i.d. The densities can either be estimated or summarized using quantiles.Instead of assuming a parametric model for the distribution (e.g. a Normal distribution with unknown expectation and variance), a non-parametric estimation assumes that the density exists and is suitably smooth (e.g. differentiable). It is then possible to estimate the unknown density function f(·).

The histogram is the oldest and most popular density estimator method. The form of the histogram depends very much on the tuning parameters class width and origin. The naive estimator was introduced by Fix and Hodges (1951) and first discussed by Rosenblatt(1956). Whittle (1958) formulated the general weight function class of estimators, while the orthogonal series estimator was introduced by Cencov (1962). The nearest neighbour estimate considered by Loftsgaarden and Quesenbury (1965), while the variable Kernel method is due to Brieman, Meisel and Purcell (1977). There are quite a few other approaches to density estimation than the kernel estimators, but the fixed bandwidth kernel estimators are used predominantly because of their simplicity.

The quantile of a distribution is defined as $Q(p)= F^{-1}(p) = inf{{x: F(x) >p}}$, $0<p< 1$, where $F(x)$ is the distribution function. Let $X_{(1)},\dots X_{(n)}$ denote the order statistics of $X_1, X_2, \dots, X_n$, and let $Qi(p)$ denote the ith sample quantile definition, then

\begin{equation}
\hat{Q_{i}(p)} = (l - \gamma)X_{(j)} + \gamma X_{(j + 1)}
\end{equation}

where,  $(j - m)/n \leq p < (j - m + 1)/n$

There have been several other nonparametric quantile estimators proposed e.g., Harrell and Davis 1982; Sheather and Marron 1990). Also sample quantiles can be of the form that are not defined for all p including hinges and other letter values (Hoaglin 1983) and related methods (Freund and Perles 1987).

@@

### Quantile based methods

@@

```{r exdot, fig.cap="10th, 20th, 30th, 40th and 50th quantiles of runs per over are drawn  across overs of the innings with no (facet 1), one (facet 2) and two (facet 3) dot balls per over. For all three cases, difference in runs between subsequent overs decrease.", fig.pos="h"}
